---
title: "Class Reference"
author: "Emmett Gartner"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

Groups loan dataframe by county of loan; then counts total # of loans
```{r}
ppp_maryland_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  )
```


To get a quick look at the top 6 rows of a data set:glimpse(wv_ppp_loans)

To store data table in righthand viewer: wv_summary<-wv_ppp_loans %>% 

When testing new code chunks, red lettering may arise as a warning — not an error. If that is the case, when you run it again, the red lettering will not appear.

mutate(new_column = value/calculation that goes into that column) In this case, we are creating a new column that takes % of loan that goes towards payroll. Can also use to update existing columns.

To turn a percentage to the 100s; percent_payroll = (payroll_proceed/amount)*100.


If else; categorizes as one thing or the other and creates new column to indicate. Code:

 in_out = if_else(
        servicing_lender_state == 'MD', "IN", "OUT"
        
Double equal signs (==) tests for equivalency.

The following code summarizes the total number of banks in and out of Maryland.
maryland_with_in_out %>% group_by(in_out)%>%summarize(count=n())

The following code tallies up how many data points fall in each month of the year and organizes them from most points to least:

maryland_ppp %>%
  mutate(Month = floor_date(date_approved, "month")) %>%
  group_by(Month) %>%
  summarise(total = n()) %>%
  arrange(desc(total))

