title: "Reverse Engineering Project"
author: "Devon Milley Timothy Dashiell Emmett Gartner"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, L.A. is slammed with record costs for legal payouts  (https://www.latimes.com/local/lanow/la-me-ln-city-payouts-20180627-story.html)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
install.packages("tidyverse")
options(scipen=999)
install.packages("janitor")
library(tidyverse)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

la_payouts <- read.csv("data/la_city_legal_payouts.csv")
View(la_payouts)

# Clean required data and prepare for analysis if needed. 

la_payouts <- la_payouts %>% mutate(AMOUNT=as.numeric(parse_number(AMOUNT)))

#sum payouts for 2018

la_payouts%>%
  group_by(Year)%>%
    summarize(
    count = n(),
    total = sum(AMOUNT))

```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: “Last budget year, the total for all types of payouts reached $201 million — the highest on record — but city officials are unsure whether that is the largest amount because records go back only two decades.”

* **General steps to reverse engineer**: 
1. Downloaded an excel sheet of all LA City Payouts between 2006-2018 from LA Times Github, deleted extraneous titles that could get in the way when converted to .csv.
2. Converted the file to a .csv.
3. Removed the /$ from the amount field.
4. Converted the amount field to a numeric datatype.
5. Grouped cases by year, summarized total payout amounts for each year.


* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-