---
title: "lab_06"
author: "Emmett Gartner"
date: "03/03/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* [install Tabula](https://tabula.technology/)
* run existing code as directed (look for **Task**).
* write code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes in the lab_06 folder and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
```

## Get Our PDF

We'll be working with the [Maryland Statistical Handbook](https://planning.maryland.gov/MSDC/Documents/md-statistical-handbook.pdf) from the state Department of Planning. It's not tiny (44MB), and you'll want to download it to a place you'll remember (like your Downloads folder).

**Task** Download the file from the above link and save it to your Downloads folder.

## Setup Tabula

**Task** Start Tabula and load the PDF, then extract a table

Start Tabula, then go to http://127.0.0.1:8080/ in your browser. Click the "Browse" button and find the Statistical Handbook file in your Downloads folder and click "open", and then click the "Import button" in Tabula. This will take a minute or two.

Let's go to page 30 of the PDF, Table 2A, "International Migration for Maryland's Jurisdictions, July 1, 2010 to July 1, 2019". This is _net migration_, meaning it shows the total number of people moving to Maryland and its regions/jurisdictions _minus_ the total number of people leaving the same, divided into 12-month periods from July to July. In Tabula, draw a box around that table's border and click the "Preview & Export Extracted Data" button. It should look pretty clean. Let's export that CSV (it should be called `tabula-md-statistical-handbook.csv` by default) to your lab_06/data folder.

## Cleaning up the data in R

Let's load it into R, and in doing so we'll skip the first two rows and add our own headers that are cleaner. `read_csv` allows us to do this ([and more](https://readr.tidyverse.org/reference/read_delim.html)):

**Task** Write code to read the CSV file into a dataframe called `international_migration`. As in the pre_lab, you may have to skip 1 or more rows and you may want to supply easier to use column names that use underscores.

```{r}
# Write the code to load the CSV file here
international_migration <- read_csv("data/tabula-md-statistical-handbook.csv", skip=2, col_names=c("State/Jurisdiction", "July_2011", "July_2012", "July_2013", "July_2014", "July_2015", "July_2016", "July_2017", "July_2018", "July_2019", "total"))%>%
view(international_migration)



```

Add a column for the type of migration ("international") and populate it:

**Task** Write the code to add a `migration_type` column to the data and give it the value of "international"

```{r}
# Write the code to add the column here
international_migration <- international_migration %>%
  mutate(migration_type = "international")

```

## Answer questions

**Q1** Which region and county/city _below_ the state level accounted for the largest percentage of international migration for the entire time period? You'll need to add and populate a column representing percent of total using `mutate`.
**A1** The Suburban Washington Region has the largest percentage of international migration for the entire time period by region, accounting for 62.8 percent of all migration below the state level. Montgomery county has the largest migration percentage of all cities and counties, with a total of 36.8 percent.

```{r}
# Write the code to produce your answer here

#create new percentage column by dividing region's total migration with Maryland's total migration, multiply x 100 for percentage share.
international_migration <- international_migration%>%
  mutate(percentage=((total)/198996)*100)%>%
        arrange(desc(percentage))%>%
          view(international_migration)
```

**Q2** Write a sentence or two that describes the data you produced in A1. Try to convey the most important idea.
**A2** The previous task determines each jurisdiction's share of international migration compared to the state of Maryland's international migrations by setting the jurisdiction's total number of migrations against Maryland's total number of migrations. The result allows us to understand the context of each jurisdiction's migrations rather than trying to interpret the total number of migrations on their own.

**Q3** Which region & jurisdiction had the largest percentage change (either negative or positive) for international migration between July 2018 and July 2017? The formula for percentage change is easy to remember: (New-Old)/Old.
**A3** The Southern Maryland Region and Calvert County had the largest percentage changes between July 2018 and July 2017, with percentages of -76.9 and -60.8, respectively.

```{r}
# Write the code to produce your answer here
international_migration <- international_migration%>%
    mutate(percentage_2018_2017=((July_2018-July_2017)/July_2017)*100)


```

**Q4** What's your best guess as to why these changes occurred, and in those places in particular?
**A4** These areas in Southern Maryland are likely less popular migration areas due to their rural nature, lack of amenities/public services, and their lack of an international community — considering their migration numbers are already quite low. Also — the immigration policies enacted by President Trump around this time likely had a profound impact on all international immigration, adding to these aforementioned factors. 


## Back to Tabula

**Task** Extract second table from Tabula

Let's go to page 31 of the PDF, Table 2B, "Domestic Migration for Maryland's Jurisdictions, July 1, 2010 to July 1, 2019". In Tabula, hit the "Clear All Selections" button and then draw a box around that table's border and click the "Preview & Export Extracted Data" button. It should look pretty clean. Let's export that CSV to your lab_06/data folder. (and rename it to `tabula-md-statistical-handbook-domestic.csv`).

## Cleaning up the data in R

Let's load it into R, and in doing so we'll skip the first two rows and add our own headers that are cleaner:

**Task** Write code to read the CSV file into a dataframe called `domestic_migration`. As in the pre_lab, you may have to skip 1 or more rows and you may want to supply easier to use column names that use underscores. Your work on the international file should serve as a guide.

```{r}
# Write the code to load the CSV file here
domestic_migration <- read_csv("data/tabula-md-statistical-handbook-domestic.csv", skip=2, col_names=c("State/Jurisdiction", "July_2011", "July_2012", "July_2013", "July_2014", "July_2015", "July_2016", "July_2017", "July_2018", "July_2019", "total"))

```

Add a column for the type of migration ("domestic") and populate it:

**Task** Write the code to add a `migration_type` column to the data and populate it with the value of "domestic"

```{r}
# Write the code to add the column here
domestic_migration <- domestic_migration %>%
  mutate(migration_type = "domestic")
```

## Answer questions
**Q5** Which Maryland individual jurisdiction saw the largest net decrease in domestic migration overall?
**A5** The Suburban Washington Region and Baltimore City saw the largest net decrease in overall domestic migration, with migrations of -107800 and -62834, respectively.

```{r}
# Write the code to produce your answer here

domestic_migration <- domestic_migration %>%
        arrange((total))%>%
          view(domestic_migration)
```

**Q6** How many regions & jurisdictions had net positive migration for July 2017, July 2018 and July 2019 (not just individually but all three together)?
**A6** 14 regions and jurisdictions 

```{r}
# Write the code to produce your answer here
domestic_migration <- domestic_migration %>%
  filter(July_2017>0, July_2018>0, July_2019>0)

```

**Q7** How would you describe this data? Is there a county or region that stands out, and why?
**A7** This data shows the most desirable/popular places for U.S. residents to move to in the state of Maryland. As a native Annapolitan, I am most surprised by the relatively high migrations to Charles County in Southern Maryland. I've visited the area and know people from it and most seem desperate to leave, rather than move there. Frederick's high migration numbers, on the other hand, do make sense to me. The town is by the foothills of the Appalachians and has a number of restaurants, stores, and boutiques. Living expenses in Frederick weren't too high then, either, although they are certainly rising now.
