---
title: "lab_05"
author: "Emmett Gartner"
date: "02/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and janitor.

**Task** load these two packages.

```{r}
# Turn off scientific notation
options(scipen=999)

library(tidyverse)
library(lubridate)
library(refinr)

```

## Load Data

You'll need to load three data sets for this:

* The West Virginia slice of the PPP loan data (lab_05.rds).
* A "lookup table" that allows you to translate NAICS (industry) numeric codes to industry titles (naics_codes.csv).
* A table of West Virginia population by county (American Community Survey, 2019 5-year averages) (wv_population_county.csv).

All three data sets are in the data folder.  Write code to load the three in the codeblock below.

**Task** Create a codeblock below this task, then read the data sets in in and assign them to appropriate variable names. There's a fourth data set you'll use in this lab, too, of selected loans in ZIP Code 25401. But there's no need to load it now.


```{r}
# Load data sets

wv_ppp_loans <- read_rds("data/lab_05.rds")

naics_codes<- read_csv("data/naics_codes.csv")

wv_population_county<- read_csv("data/wv_population_county.csv")
  



```

## Answer questions

**Q1.** In the data folder, there is a csv called zip_25401_loan_sample.csv.  It contains a sample of loans from West Virginia ZIP Code 25401.

As we read earlier this semester, [multiple loan applications coming from multiple businesses at the same residential street address](https://www.nytimes.com/2021/08/17/business/ppp-fraud-covid.html) might point to fraud. Or it could alert us to companies that used [multiple corporate entities to get more money than envisioned](https://www.propublica.org/article/different-names-same-address-how-big-businesses-got-government-loans-meant-for-small-businesses) by the original law that authorized the program.   

You are going to examine this data to see if we can find a lot of loans coming from the same address.  Here's the problem: the street address field is pretty messy.  The same address appears with minor variations --  "1003 Sushruta Dr" vs "1003 SUSHRUTA DR" -- that will prevent proper grouping.

First, upload the data into Open Refine and standardize/clean the address field - make sure to make a copy of that column to work on and to try all of the options for clustering. If you've done it properly, you should have 65 discrete addresses.

Then export the data from Open Refine, and move it to the proper folder.

Next, load it in the codeblock below, assigning it to an appropriate variable name.

Then answer these questions:
* What is the street address in this data that has the most loans?
* How many loans are there at that street address?
* What are the names of the businesses at that address?

**A1.**
126 E Burke St has the most loans, with 6 total loans. The names of the businesses at that address are: DREW HOLDINGS LLC, BRIX27, LLC, ABRAHAM ASHTON, HUB CO-OP LLC, HUB CO-OP, LLC, and RONIN PROPERTIES LLC.

```{r}

#load cleaned file for West Virginia ZIP Code 25401

zip_25401_loan_sample<- read_csv("data/zip-25401-loan-sample-csv.csv")

#street address with the most loans

zip_25401_loan_sample%>%
  group_by(clean_address) %>%
    summarise(
    count=n()) %>%
    arrange(desc(count))

#Business names at  126 E Burke St

zip_25401_loan_sample%>%
  filter(clean_address == "126 E Burke St")



```

**Q2.** Do some web research on the businesses that answered question 1.  

Google the street address.  Google the business names and search for their corporate records on [Open Corporates](https://opencorporates.com/). Be sure to find the website of the name of the company that appears twice in the list of businesses and develop an understanding of what it does.

Based on your research, does it seem suspicious that this collection of businesses all got loans using the same address? Why or why not. Use specific facts identified in your research to support your position.

**A2.**
Yes, the loan behavior of this collection of businesses does seem a little suspicious. Five of the listed businesses (excluding the duplicate HUB CO-OP LLC) have the same name listed as a Director/Officer, Abraham Ashton. There is also another name, Robert Johnson, that is listed at a leadership position in these businesses. 

**Q3.** Start by using the West Virginia slice of the PPP loan data that you loaded at the start of the lab to create a subset of PPP loans in West Virginia's second largest county (which you can find in wv_population_county). And then use that table you created to answer the following questions:

* Which city in that county had the highest number of loans?
* In that city, which industry title had more loans than any other industry title?

```{r}
berkeley_ppp_with_naics <- cleaned_berkeley_county_ppp.csv %>% left_join(naics_codes, by="naics_code")%>%
  mutate(zip5 = str_sub(zip, 1, 5))
```


Requirement: you MUST export a dataframe of PPP loans from R Studio at some point in the process (not necessarily at the beginning!), load it into Open Refine, clean the city column, export it from Open Refine, and reimport into R Studio. To export data from R Studio, you will use the write_csv() function, which works like this if we're using the wv_population_county dataframe:

```{r}
# write_csv accepts 2 arguments: the dataframe and the path to the file you write out
write_csv(wv_population_county, "data/wv_population_county.csv")

#filter berkeley county ppp loans
berkeley_county_ppp <- wv_ppp_loans%>%
  filter(project_county_name == "BERKELEY")

#write berkeley_county_ppp as .csv file, export to OpenRefine and clean city name data
write_csv(berkeley_county_ppp, "data/berkeley_county_ppp.csv")

#load cleaned berkeley county data

cleaned_berkeley_county_ppp.csv<- read_csv("data/berkeley-county-ppp-csv.csv")

#filter to determine city with the most loans in berkeley county

cleaned_berkeley_county_ppp.csv%>%
    group_by(cleaned_city)%>%
    summarise(
    count=n()) %>%
    arrange(desc(count))

#filter to Martinsburg and determine industry with the most loans
berkeley_ppp_with_naics%>%
    filter(cleaned_city == 'Martinsburg')%>%
        group_by(title)%>%
        summarise(
        count=n()) %>%
        arrange(desc(count))

```

Guidance: there are a lot of steps you'll need to take to answer this question. You may find it helpful to write out in English what you plan to do step-by-step before you start writing code.   

**A3.**
Martinsburg had the most loans, with a total of 1166. The industry with the most loans has a naics code of 722511 and is defined as "full-serivce restaurants." It accounted for 50 loans in Martinsburg.

**Q4.** What are your two best hypotheses, which you'd need to confirm with reporting and further analysis, that explain why that industry is at the top of the list?

**A4.**
According to a PPP breakdown by finder.com, the Accommodation and Food Services industry received the most total loan dollars from the PPP across the United States. This makes sense that "full-service restaurants" would be the industry with the most loans in Martinsburg, WV. Additionally, because Martinsburg has a population of 18,777 according to the 2020 census, it does not appear to be the hub of any outside industries — making the restaurant business a dominant force in town.

**Q5.** Start with a dataframe of loans to all businesses in the city and industry that answered question 3. Answer the following questions:
* What is the name of the business that got the highest approved loan amount?
* How much was it for?
* When was it approved?
* How many jobs does the data say were retained?
* Is there a difference between the business' name in the PPP data and the name its customers know it by? If so, what is that name?
* How many locations does this business have?
* Did one of its locations close during the pandemic, either before or after it got the loan?

```{r}
#filter to naics_code of 722511 in Martinsburg, WV
martinsburg_restaurants_ppp_naics <- berkeley_ppp_with_naics%>%
    filter(cleaned_city == 'Martinsburg', title == 'Full-Service Restaurants')

#arrange businesses by loan sums, highest to lowest amounts
martinsburg_restaurants_ppp %>%
    group_by(name) %>%
    arrange(desc(current_approval_amount))



```

Hint: you will not find the answers to the last three questions in the data.  You could call them directly to get that information, but I don't want you to do that for this assignment.  Instead, do some web research. I would start by Googling the company name from the data and looking at the page that comes up for the business from at http://apps.sos.wv.gov/. I would use information I found on that page and use info about the company from Google, the [Wayback machine](https://archive.org/web/) (which lets you look at older versions of a company's website), Yelp, and Facebook.

**A5.**
COTTLE CHRISTI L LLC had the highest approved loan amount of all Martinsburg restaurants with a total of \$280434.00. It was approved on 2021-02-17 and claims to have retained 94 jobs. 

Following answers found at this website: https://apps.sos.wv.gov/business/corporations/organization.aspx?org=338507

Yes, there is a difference. The aforementioned LLC owns the following brick and mortar restaurants: KITZIE'S CAFE, KITZIE'S CAFE II, KITZIE'S OF INWOOD, KITZIE'S OF SPRING MILLS, KITZIE'S RESTUARANT & LOUNGE and RIVERBEND BAR & GRILL. Six restaurants total. 

KITZIE'S OF INWOOD is closed, and after using Wayback Machine to look at the website for the Kitzie's chain, it appears that the Inwood location was taken down from the website around June 2019, before receiving the loan. I had trouble tracking down Kitzie's Cafe and Kitzie's Cafe II — there is no trace of their existence on the internet, despite being listed on the West Virginia Secretary of State website.