---
title: "lab_04"
author: "Emmett Gartner"
date: "02/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
```

## Load Data

Let's keep working with Maryland loan applications contained in the PPP data.

**Task** Read in the lab_04.rds file and assign it to the variable `maryland_ppp`.

```{r}
# add code here to accomplish the task

maryland_ppp <- read_rds('lab_04.rds')


```

To answer the questions below, you'll need to join both the NAICS code data and the ZCTA demographic data from the pre-lab. There's a particular set of steps you'll need to perform to make that work properly, so take your time and check your work.

First, let's load the NAICS code CSV file we have in the data folder:

**Task** Read in the naics_codes.csv file and assign it to the variable `naics_codes`.

```{r}
# add code here to accomplish the task

naics_codes <- read.csv('data/naics_codes.csv')

```

Next join the two and then create a 5-character `zip5` column that we can use to join to the ZCTA data:

**Task** Write code to join the maryland_ppp and naics_codes data, assigning the result to a new variable. Then add a new `zip5` column.

```{r}
# Create new data frame, join by naics_codes, create new column titled "zip5"


maryland_ppp_with_naics <- maryland_ppp %>% left_join(naics_codes, by="naics_code")%>%
  mutate(zip5 = str_sub(zip, 1, 5))

```

Let's add the Zip Code Tabulation Area dataset. You'll need to make the `ZCTA5N` column a character column, not a numeric one.

**Task** Read in the Zip Code Tabulation Area dataset in the data folder and change the datatype of the `ZCTA5N` column to character. IMPORTANT: you can see what the column names represent by checking the data documentation, which can be found here: https://geodata.md.gov/imap/rest/services/Demographics/MD_CensusData/FeatureServer/1. You will need to refer to this to answer the questions below.

```{r}
# Read ZCTA file
maryland_zcta<- read.csv('data/maryland_zcta.csv')%>%

# Edit data of ZCTA5N, set to character
mutate(across(ZCTA5N, as.character))

#Check previous task
glimpse(maryland_zcta)
```

Now we can join the PPP data to the ZCTA data:

**Task** Join the Maryland PPP data (with NAICS already joined) with the ZCTA data, specifying the join column in both dataframes.

```{r}
# Join Maryland PPP with ZCTA

maryland_ppp_with_naics_and_demographics <- maryland_ppp_with_naics %>% left_join(maryland_zcta, by=c("zip5"="ZCTA5N"))

```

## Answer questions

**Q1** What are the top 5 industries (using the NAICS title) with the most PPP applications in Maryland?
**A1** The top 5 industries are: Taxi Service, Beauty Salons, Full-Service Restaurants, Offices of Real Estate Agents and Brokers, and All Other Personal Services.

```{r}
# Arrange top 5 industries w/most PPP applications by NAICS title
  maryland_ppp_with_naics_and_demographics %>%
  
  group_by(title) %>%
    summarise(
    count=n()) %>%
    arrange(desc(count))


```

**Q2** What are the top industries (using the NAICS title) with the most PPP applications in Maryland where the loan amount is more than $150,000?
**A2** The top 5 industries are: Full-Service Restaurants, Offices of Physicians (except Mental Health Specialists), Plumbing, Heating, and Air-Conditioning Contractors, Offices of Dentists, and Commercial and Institutional Building Construction.

```{r}
# Write the code to produce your answer here

  maryland_ppp_with_naics_and_demographics %>%
  filter(amount > 150000)%>%
  group_by(title) %>%
    summarise(count=n()) %>%
    arrange(desc(count))

```

**Q3** Which Maryland zip code with at least 20 percent non-Hispanic Asian population had the largest sum of PPP loans? Also show the number (count) of loans.
**A3** The zip code 20850 had the largest sum of PPP loans, a total value of \$474525058. The count for 20850's loans is 3186.
```{r}
# Maryland zip code with at least 20 percent with population non-Hispanic Asian population
  maryland_ppp_with_naics_and_demographics %>%
  filter(PNHA > 20)%>%
  group_by(zip5) %>%
    summarise(total = sum(amount)) %>%
    arrange(desc(total))
    
# Count loans in zip 20850
maryland_ppp_with_naics_and_demographics %>%
    group_by(zip5 == '20850')%>%
    summarize(
    count = n())

#Check first task
  maryland_ppp_with_naics_and_demographics %>%
   group_by(zip5) %>%
    summarise(total = sum(amount)) %>%
    arrange(desc(total))

```

**Q4** Which Maryland zip code had the largest sum of loans where the reported race of the applicant was "Asian", and what is the percentage of non-Hispanic Asian population in that zip code? Also show the number (count) of loans.
**A4** The zip code is 20850 and the PNHA percentage is 23.5.

```{r}
# Zip code with largest sum of loans where race of applicant was "Asian", arranged in descending totals.
    maryland_ppp_with_naics_and_demographics %>%
      filter(race == 'Asian')%>%
        group_by(zip5)%>%
          summarise(total = sum(amount))%>%
                arrange(desc(total)) 
# Count
    maryland_ppp_with_naics_and_demographics %>%
        group_by(zip5 == '20850')%>%
        summarize(
        count = n())

```

**Q5** What might explain why 20850 has a significant Asian population (23.5%) and only 240 of the 3,186 loans (7.5%) are to Asian applicants?
**A5** When we previously scrutinized this data, we determined that a large number of applicants omitted the "race" portion of the PPP application. It is likely that this discrepancy originates from this frequent omission.


**Q6** Check to see if any Maryland zip codes that have loans where the PPP data lists the applicant as being Asian BUT are located in a zip code that does not appear to have Asian population according to the ZCTA data. How many loans do they have and what is the total amount? Take a look at these zip codes - what could explain this?

**A6** There are three zip codes that meet these parameters and they have a total of 13 loans. The total loan amount is \$7253408. This could be explained by the fact that the owner of the business resides outside of the zip code that their business is in.

```{r}
# Tasks: filter loans by race == 'Asian' and PNHA is 0, then count and summarize by zip.
  
  maryland_ppp_with_naics_and_demographics %>%
  filter(race == 'Asian', PNHA == 0)%>%
  group_by(zip5)%>%
    summarize(
    count = n(),
    total = sum(amount))

#  Find single total loan amount that meets these parameters
    maryland_ppp_with_naics_and_demographics %>%
      filter(race == 'Asian', PNHA == 0)%>%
          group_by(zip5)%>%
            summarise(total =sum(amount))%>%
              summarise(total = sum (total))

```

**Q7** How many industries have exactly one loan in the data (hint: you can filter _after_ using `summarise` on any totals you create). From a newsworthiness standpoint, which of these is most worthy of further exploration, and why?

**A7** 44 total industries with exactly one loan. I am curious about the Newsprint Mills industry only having one loan, given the troubled economic state of newspapers and print media in general. You would think that any and all print mills would be after free government money.

```{r}
# Filter to industries w/1 loan.
  maryland_ppp_with_naics_and_demographics %>%
    group_by(title)%>%
     summarise(count=n())%>%
      filter(count==1)



```
